---
title: Lab Eight -- Probability # Title of the document

# Bibliography Details
bibliography: Lab8References.bib  # Name of bib file in the same folder 
bibliographystyle: apa            # Bibliography style

execute:               # Default settings for code blocks
  warning: false       # Do not print warnings
  message: false       # Do not print messages
  size: "scriptsize"   # Set code size to script size

# Below will do html by default. To render the entire project
# Use the following code:
# quarto::quarto_render("Lab8WriteUp.qmd")
format:
  pdf:        # Specify .pdf formatting options 
    documentclass: article     # Document Class
    number-sections: true      # Number Sections
    reference-section-title: "References" # Title for reference section
    code-line-numbers: true    # Line numbers printed in code
    cap-location: top          # Captions at the top of tables 
    geometry:
      - top=0.75in
      - left=0.75in
      - bottom=0.75in
      - right=0.75in
  html:       # Specify HTML formatting options 
    number-sections: true      # Number Sections
    code-fold: true            # Button to show code
    code-line-numbers: true    # Line numbers printed in code
    cap-location: top          # Captions at the top of tables and figures
format-links: [pdf]
---
```{r}
#| echo: false

default_chunk_hook  <- knitr::knit_hooks$get("chunk")

latex_font_size <- c("Huge", "huge", "LARGE", "Large", 
                     "large", "normalsize", "small", 
                     "footnotesize", "scriptsize", "tiny")

knitr::knit_hooks$set(chunk = function(x, options) {
  x <- default_chunk_hook(x, options)
  if(options$size %in% latex_font_size) {
    paste0("\n \\", options$size, "\n\n", 
      x, 
      "\n\n \\normalsize"
    )
  } else {
    x
  }
})
```

```{r}
#| label: load-packages
#| include: false
library(tidyverse)
library(patchwork)
library(knitr)
library(extraDistr)
```

**Tips**

* Complete the tasks below. Make sure to start your solutions in on a new line that starts with "**Solution**:".
* Make sure to use the Quarto Cheatsheet. This will make completing and writing up the lab *much* easier.
* Make sure to use "enter" to make your code readable, so it doesn't run off the page. I switched the Quarto document to automatically render to .pdf, so you can see what you're submitting by default. You can switch back for the highlighting by moving the html block above the pdf block in the YAML header.
* Don't forget that you can copy-and-paste code when you're doing something similar as we did in class!
* Make sure to plot all computed probabilities. This is good `ggplot` practice AND will help make sure you don't make a mistake when computing the probability.

# The Binomial and Negative Binomial Distributions

## The Binomial Distribution
In many introductory courses, we make the simplifying assumption of independence in Bernoulli trials. For example, when playing best two-out-of-three legs in a darts match, we would assume Player A has an 70% chance of winning each leg. We write:
$$X \sim \text{Binomial}(n=3, p=0.70)$$
where

* $X$ is the number of legs won
* $n$ is the number of trials 
* $p$ is the probability of the event of interest

Of note, it is not necessarily the case that the players play all three legs -- if either player wins legs one and two the match ends.

There are eight possible outcomes for Player A:

* (win, win, "win") $\longrightarrow$ Player A wins
* (win, win, "lose") $\longrightarrow$ Player A wins
* (win, lose, win) $\longrightarrow$ Player A wins
* (win, lose, lose) $\longrightarrow$ Player A loses
* (lose, win, win)  $\longrightarrow$ Player A wins
* (lose, win, lose) $\longrightarrow$ Player A loses
* (lose, lose, "win") $\longrightarrow$ Player A loses
* (lose, lose, "lose") $\longrightarrow$ Player A loses

**Note**: I use quotations above to denote outcomes where the last match isn't actually played.

**Use the Binomial distribution to compute the probability that Player A wins the match.  Make sure write out your probability statement and work to write it in terms of the distribution function (PMF/CDF) to plot the probability using `ggplot`.**

**Solution**

$P(X>=2) = 1 - F(1) = f(2) + f(3)$
```{r}
bin.cdf = 1- pbinom(1, 3, 0.7) 
bin.pmf = dbinom(2, 3, 0.7) + dbinom(3,3, 0.7)
bin.cdf
bin.pmf
```

```{r}
#| label: Binom.PMF/CDF
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize
ggdat <- tibble(x = 0:3) |>
 mutate(PMF = dbinom(x, 3, 0.7), 
        CDF = pbinom(x, 3, 0.7))
#make data contain both PMF and CDF to be specified later
#use the _binom function for the binomial distribution

PMF.plot <- ggplot() +
 geom_linerange(data = ggdat,
aes(x = x, ymin = 0, ymax = PMF)) + #Specifies PMF data
geom_hline(yintercept = 0) +
scale_x_continuous("x", breaks=0:3) + #0:3 for diff outcomes
ylab("Probability") +
theme_bw()

CDF.plot <- ggplot() +
  geom_step(data=ggdat, aes(x = x, y = CDF)) + #ensures its not a smooth line
  geom_point(data=ggdat, aes(x = x, y = CDF)) + #gives points
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks=0:3) + #ensures 4 possible outcomes
  ylab("Cumulative Probability") +
  theme_bw()

PMF.plot | CDF.plot
```

```{r}
#| label: Binom.Solution
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggdat <- tibble(x = 0:3) |>
mutate(PMF = dbinom(x, prob = 0.70, size = 3),
      CDF = pbinom(x, 3, 0.7))


PMF.plot <- ggplot() +
  geom_linerange(data = ggdat,
  aes(x = x, ymin = 0, ymax = PMF)) +
geom_linerange(data = ggdat |> filter(x >= 2), # when x is greater than 2
  aes(x = x, ymin = 0, ymax = PMF), 
  color = "darkorchid2", linewidth = 2, alpha = 0.50) + #highlight w color
geom_hline(yintercept = 0) +
scale_x_continuous("x", breaks = 0:3) +
ylim(0, 1) +
ylab("Probability") +
theme_bw()

CDF.plot <- ggplot() +
  geom_step(data=ggdat, aes(x = x, y = CDF)) +
  geom_point(data=ggdat, aes(x = x, y = CDF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks=0:3) +
  ylab("Cumulative Probability") +
  theme_bw()

PMF.plot | CDF.plot
```

## The Negative Binomial Distribution
In intermediate courses, we learn the Negative Binomial distribution that we use to model the number of "failures" until the $r$th success. Note that this approach still makes the simplifying assumption of independence in Bernoulli trials.

**Specify the two cases necessary to compute the probability. Fully explain these probability statements in the context of a best two-out-of-three match and the Negative Binomial distribution. Finally, compute the probability that Player A wins the match and show you get the same result as the Binomial approach. Make sure write out your probability statement and work to write it in terms of the distribution function (PMF/CDF) to plot the probability using `ggplot`.**

**Solution**

In order to compute the probability with a negative binomial distribution, we need to restrict $x$ to $0:2$ because if there are more than two failures, the match will end. Additionally, we need the $r$th success to be two, as that is our event of interest (player A winning). 
To calculate this probability, we can use:

$P(X<=1) = f(1) + f(2) = F(1)$

```{r}
dnbinom(0,2,.7) + dnbinom(1,2,.7)
pnbinom(1, 2, 0.7)
```

```{r}
#| label: Neg.cdf/pmf
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggdat <- tibble(x = 0:2) |>
mutate(PMF = dnbinom(x, size = 2, prob = 0.70), #2 for reasons above
       CDF = pnbinom(x, size = 2, prob = 0.70))

 PMF.plot <- ggplot() +
 geom_linerange(data = ggdat,
 aes(x = x, ymin = 0, ymax = PMF)) +
 geom_hline(yintercept = 0) +
scale_x_continuous("x", breaks = 0:2) +
 ylab("Probability") +
theme_bw()

CDF.plot <- ggplot() +
geom_step(data = ggdat, aes(x = x, y = CDF)) +
geom_point(data = ggdat, aes(x = x, y = CDF)) +
geom_hline(yintercept = 0) +
scale_x_continuous("x", breaks = 0:2) +
ylab("Cumulative Probability") +
theme_bw()

PMF.plot | CDF.plot
```

```{r}
#| label: Neg.Win
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggdat <- tibble(x = 0:2) |>
mutate(PMF = dnbinom(x, size = 2, prob = 0.70),
       CDF = pnbinom(x, size = 2, prob = 0.70))

 PMF.plot <- ggplot() +
 geom_linerange(data = ggdat,
            aes(x = x, ymin = 0, ymax = PMF)) +
 geom_linerange(data = ggdat |> filter(x < 2), #when x < 2 highlight lines
            aes(x = x, ymin = 0, ymax = PMF),
            color = "darkorchid2", linewidth = 2, alpha = 0.50) +
 geom_hline(yintercept = 0) +
scale_x_continuous("x", breaks = 0:2) +
 ylab("Probability") +
theme_bw()

CDF.plot <- ggplot() +
geom_step(data = ggdat, aes(x = x, y = CDF)) +
geom_point(data = ggdat, aes(x = x, y = CDF)) +
  geom_segment(aes(x = 1, y = 0.784, xend = 0, yend = 0.784),
               color = "darkorchid2", linetype = "dashed", size = 1) +
    geom_point(aes(x = 1, y = 0.784), color = "darkorchid2", size = 2)+
geom_hline(yintercept = 0) +
scale_x_continuous("x", breaks = 0:2) +
ylab("Cumulative Probability") +
theme_bw()
#will draw line from x=1 to probability and highlight dot
PMF.plot | CDF.plot
```

# The Beta Binomial Distribution
One approach to generalize this modeling is to use a Beta-Binomial distribution. This distribution arrises when we let $p$ be a random variable that follows a Beta distribution:
\begin{align*}
  P &\sim \text{Beta}(\alpha, \beta)\\
  X &\sim \text{Binomial}(n, P).
\end{align*}
This distribution is not part of base `R`, but is available using the `_bbinom` functions from the `extraDistr` package [@extraDistr].

The primary benefit is that the Beta-Binomial enables us to specify some uncertainty about the skill of Player A, as $p$ does not have to be fixed (e.g., $p=0.70$). Suppose the result of the best two-out-of-three match follows a Beta-Binomial distribution  where $\alpha=7$ and $\beta=3$.

## The Beta Part
1. **Plot the corresponding Beta distribution. Interpret the plot in the context of a best two-out-of-three match and the Binomial distribution.**\
**Hint**: The Beta distribution is available in base `R` via `_beta()`.

**Solution**

```{r}
#| label: Beta.pdf/cdf
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

alpha = 7
beta = 3

ggdat <- tibble(x = seq(0, 1, length.out=100)) |>
mutate(PDF = dbeta(x, alpha, beta), #using _beta for beta distribution
       CDF = pbeta(x, alpha, beta))

PDF.plot <- ggplot(ggdat) +
geom_line(aes(x = x, y = PDF)) +
geom_hline(yintercept = 0) +
xlab("p") + # p label bc this is for a prob not number of events
ylab("Density") +
theme_bw()

CDF.plot <- ggplot(ggdat) +
geom_line(aes(x = x, y = CDF)) +
geom_hline(yintercept = 0) +
xlab("p") +
ylab("Cumulative Probability") +
theme_bw()
#geom_line instead of geom_linerange for continuous distribution
PDF.plot | CDF.plot
```

```{r}
#| label: Beta.Win
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

alpha = 7
beta = 3

ggdat <- tibble(x = seq(0, 1, length.out=100)) |>
mutate(PDF = dbeta(x, alpha, beta),
       CDF = pbeta(x, alpha, beta))

PDF.plot <- ggplot(ggdat) +
  geom_line(aes(x = x, y = PDF)) +
  geom_ribbon(data = ggdat|> filter(x >=0.5), #fills area after x= 0.5
  aes(x=x, ymin=0, ymax= PDF), #max keeps it from going above line
  fill= "darkorchid2", alpha= .5) +
  geom_hline(yintercept = 0) +
  xlab("p") +
  ylab("Density") +
  theme_bw()

CDF.plot <- ggplot(ggdat) +
geom_line(aes(x = x, y = CDF)) +
geom_point(data=tibble(x=0.5) |> mutate(CDF = pbeta(x, 7, 3)),
       aes(x = x, y=CDF), color="darkorchid2")+ #makes visible dot to mark p=0.5
geom_hline(yintercept = 0) +
xlab("p") +
ylab("Cumulative Probability") +
theme_bw()

PDF.plot | CDF.plot
```

These plots portray the probability distributions of the probability of player A winning a match. For example, the probability of the probability of player A winning at most 75% of the time is about 60%. In terms of the binomial distribution, this inconsistent probability in player A winning each round makes each win less certain. Ultimately, the Beta distribution allows a greater variation in the probability of Player A to win each match in the binomial distribution.

1. How does the Beta incorporate uncertainty about the skill of Player A? **Compute the mean and variance of the distribution.**\
**Hint**: You will find the `integrate(...)` function helpful here. 
\begin{align*}
  E(P) &= \mu_P = \int_{0}^1 p f_P(p) dp\\
  var(P) &= \sigma^2_P = \int_{0}^1 (p-\mu_P)^2 f_P(p) dp\\
\end{align*}

**Solution**

```{r}
p= seq(0, 1, length.out=100) # prob can only be 0 to 1
fofp = function(p){p * dbeta(p, 7, 3)} 
#function of p with p times the distribution
mean = integrate(fofp, 0, 1)
mean

varP = function(p){(p- mean$value)^2 * dbeta(p,7,3)}
#mean$value ensures error isn't returned too
integrate(varP, 0, 1)
```

1. **What is the probability that Player A has a win probability of exactly 0.70?  Make sure write out your probability statement and work to write it in terms of the distribution function (PDF/CDF) to plot the probability using `ggplot`.**\
**Hint:** Think carefully about the properties of continuous distributions here.

**Solution**

Player A will not have a win probability of exactly any number because a continuous distribution uses area to calculate the probability, and the area of a line (a single value) is 0. 

1. **What is the probability that Player A has a win probability of at least 0.50?  Make sure write out your probability statement and work to write it in terms of the distribution function (PDF/CDF) to plot the probability using `ggplot`.**

**Solution**

$P(p >= 0.5) = 1 - F(p <= 0.5)$

```{r}
1 - pbeta(.5, 7, 3)
```

The probability that player A has a win probability of at least .5 is approximately 0.910. I found this using the CDF where the x value is 0.5 because it gives the proability of getting at most 0.5. By subtracting this probability from 1, I could get the probability that the winning probability is at least 0.5. Because a single winning probability has a probability of 0, I did not have to factor in the winning probability of 0.5.

1. **What is the 90th percentile of Player A's win probability based on this Beta distribution? Make sure write out your probability statement and work to write it in terms of the distribution function (PDF/CDF) to plot the percentile using `ggplot`.**

**Solution**

$P(X <= 0.9) = F(p = 0.9)$

```{r}
qbeta(.9, 7, 3)
```


The 90th percentile of Player A's win probability is about 0.871.
```{r}
ggdat <- tibble(x = seq(0, 1, length.out=1000))|>
mutate(PDF = dbeta(x, 7, 3, log = FALSE),
CDF = pbeta(x, 7, 3, lower.tail = TRUE, log.p = FALSE))

PDF.plot <- ggplot() +
  geom_line(data = ggdat,
        aes(x = x, y = PDF)) +
  geom_ribbon(data = ggdat|> filter(x <=0.9), #fills area before x= 0.9
          aes(x=x, ymin=0, ymax= PDF), #max keeps it from going above line
          fill= "darkorchid2", alpha= .5) +
  geom_hline(yintercept = 0) +
  ylab("Density") +
  xlab("p")+
  theme_bw()

CDF.plot <- ggplot() +
  geom_line(data = ggdat, aes(x = x, y = CDF)) +
  geom_point(data=tibble(x=0.9) |> mutate(CDF = pbeta(0.9, 7, 3)),
       aes(x = x, y=CDF), color="darkorchid2")+ 
        #makes visible dot to mark p=0.9
  geom_hline(yintercept = 0) +
  ylab("Cumulative Probability") +
  xlab("p")+
  theme_bw()

PDF.plot | CDF.plot
```

## The Beta-Binomial Part
**Compute the probability that Player A wins the match. Compare your result to the Binomial and Negative Binomial results. Make sure write out your probability statement and work to write it in terms of the distribution function (PMF/CDF) to plot the probability using `ggplot`**

```{r}
library("extraDistr")
```

```{r}
ggdat <- tibble(x = 0:3) |>
mutate(PMF = dbbinom(x,3,7,3),
      CDF= pbbinom(x,3,7,3))
#alpha, beta, and number of games is fixed, wins is between 0 and 3

PMF.plot <- ggplot() +
  geom_linerange(data = ggdat,
    aes(x = x, ymin = 0, ymax = PMF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylab("Probability") +
  theme_bw()

CDF.plot <- ggplot() +
  geom_step(data = ggdat, aes(x = x, y = CDF)) +
  geom_point(data = ggdat, aes(x = x, y = CDF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylab("Cumulative Probability") +
  theme_bw()

PMF.plot | CDF.plot
```

$P(X >= 2) = F( X <= 1)$

```{r}
1 - pbbinom(q = 1, size = 3, alpha = 7, beta = 3) 
# probability of everything but 1 and 0 successes unfer this distributions
```

```{r}
ggdat <- tibble(x = 0:3) |>
mutate(PMF = dbbinom(x,3,7,3),
       CDF= pbbinom(x,3,7,3))

PMF.plot <- ggplot() +
  geom_linerange(data = ggdat,
  aes(x = x, ymin = 0, ymax = PMF)) +
  geom_linerange(data = ggdat |> filter(x >= 2), 
  #highlights when x is at least 2
    aes(x = x, ymin = 0, ymax = PMF),
  color = "darkorchid2", linewidth = 2, alpha = 0.50) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylab("Probability") +
  theme_bw()

CDF.plot <- ggplot() +
  geom_step(data = ggdat, aes(x = x, y = CDF)) +
  geom_point(data = ggdat, aes(x = x, y = CDF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylab("Cumulative Probability") +
  theme_bw()

PMF.plot | CDF.plot
```

Unlike the binomial probability, there is a greater chance of winning all three matches, demonstrating that this distribution factors in the possibility of a "hot streak."

## The Why
Check the expected value and variance of each distribution.

Analytically, we have
\begin{align*}
  E(X) &= np  &&var(X) = np(1-p) \tag{Binomial Distribution}\\
  E(X) &= n\left(\frac{\alpha}{\alpha+\beta}\right)  &&var(X) = \frac{n \alpha \beta (\alpha + \beta + n)}{(\alpha+\beta)^2(\alpha + \beta + 1)} \tag{Beta-Binomial Distribution}
\end{align*}

**For each distribution, compute the mean number of legs won for player A.**\
**Hint**: You can perform the sum over the finite support set here. 
\begin{align*}
  E(X)   = \mu_X    &= \sum_{i=0}^3 i f_X(i)\\
  var(X) = \sigma^2_X &= \sum_{i=0}^3 (i-\mu_X)^2 f_X(i)
\end{align*}

**Solution**

```{r}
#just making R do arithmetic
Binomial.Mean = 3*.7
Binomial.Variance = 3*.7*(1-.7) 
Binomial.Mean
Binomial.Variance
Beta.Binomial.Mean = 3*(7/(7+3))
Beta.Binomial.Variance = ((3*7*3)*(7+3+3)/(((7+3)^2)*(3+7+1)))
Beta.Binomial.Mean
Beta.Binomial.Variance
```

```{r}
#performing the same calculations as above, but with built in functions

binmean = dbinom(1,3, .7) + 
          2 * dbinom(2, 3, .7) + 
          3 * dbinom(3, 3, .7) + 
          0 * dbinom(0, 3, .7)

binmean

binvariance = ((0-binmean)^2)*(dbinom(0,3,.7)) + 
              ((1-binmean)^2)*(dbinom(1,3,.7)) + 
              ((2-binmean)^2)*(dbinom(2,3,.7)) + 
              ((3-binmean)^2)*(dbinom(3,3,.7))

binvariance

betamean = dbbinom(x = 1, size = 3, alpha = 7, beta = 3) + 
          2 * dbbinom(x = 2, size = 3, alpha = 7, beta = 3) + 
          3 * dbbinom(x = 3, size = 3, alpha = 7, beta = 3) + 
          0 * dbbinom(x = 0, size = 3, alpha = 7, beta = 3)

betamean

betavariance = ((0-betamean)^2)*(dbbinom(x = 0, size = 3, alpha = 7, beta = 3)) + 
               ((1-betamean)^2)*(dbbinom(x = 1, size = 3, alpha = 7, beta = 3)) + 
               ((2-betamean)^2)*(dbbinom(x = 2, size = 3, alpha = 7, beta = 3)) + 
               ((3-betamean)^2)*(dbbinom(x = 3, size = 3, alpha = 7, beta = 3))

betavariance
```

# Altham's Multiplicative Binomial Distribution
While the Beta-Binomial enables us to model cases where winning has a clustering effect (e.g., a "hot hand").

## Probability Mass Function
The mass function for this distribution is below. While it used to be available in the `MBD` package for `R`, this package is no longer available.
$$f_X(x) = \frac{1}{c} {n \choose x} p^x (1-p)^{n-x} \theta^{x(n-x)}I(x \in \{0, 1, ..., n\})$$
where
$$c = \sum_{i=0}^n {n \choose i} p^i (1-p)^{n-i} \theta^{i(n-i)}$$ 

**Write a function in `R` called `daltbinom()` for computing the PMF of this distribution. Use your function to plot the distribution where $p=0.70$ where $\theta=0.50$, $\theta=1$, and $\theta=1.50$. Comment on the differences in the context of a best two-out-of-three match**

**Solution**

```{r}
#| label: Alt.PMF
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

daltbinom = function(x, n, p, t){
  i = 0:n 
  c = sum(choose(n, i)*(p^i)*((1-p)^(n-i)*(t^(i*(n-i)))))
  pdf= (1/c)* choose(n, x)*(p^x)*((1-p)^(n-x)*(t^(x*(n-x))))
}
# in the new daltbinom function of x, n, p, and t
# for i is between 0 and n,
#  sum the binomial coefficent of n, i, and the following arithmetic
# let this be c
# the pdf equals 1/c times the binomial coefficient of n and x and the following arithmetic

ggdat <- tibble(x = 0:3) |>
  mutate(PMF = daltbinom(x,3,0.70,0.5)) 
#restricts data to previous function and sets n, p, and theta(0.5) as fixed

PMF.plot.1 <- ggplot(ggdat) +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("X Where Theta = 0.5", breaks = 0:3) +
  ylim(0,1)+ #prevents anything from being above 1
  ylab("Probability") +
  theme_bw()

ggdat <- tibble(x = 0:3) |>
  mutate(PMF = daltbinom(x,3,0.70,1))
#restricts data to previous function and sets n, p, and theta (1) as fixed

PMF.plot.2 <- ggplot(ggdat) +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("X Where Theta = 1", breaks = 0:3) +
  ylim(0,1)+
  ylab("Probability") +
  theme_bw()

ggdat <- tibble(x = 0:3) |>
  mutate(PMF = daltbinom(x,3,0.70,1.5))
#restricts data to previous function and sets n, p, and theta(1.5) as fixed

PMF.plot.3 <- ggplot(ggdat) +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("X Where Theta = 1.5", breaks = 0:3) +
  ylim(0,1)+
  ylab("Probability") +
  theme_bw()

PMF.plot.1 | PMF.plot.2 | PMF.plot.3

#next q # F(x)=f(0)+ f(1) + ... + f(x)
```

This demonstrates that as theta increases, there is a greater chance of winning two out of three of the matches rather than all three. Additionally, when theta equals 1, the distribution is just a normal binomial probability distribution where probability is independent of each trial. So as theta moves away from one, the probability of winning a match becomes less independent of the past match outcomes.

## Cumulative Distribution Function
**Use your `daltbinom()` function to create a `paltbinom()` function for computing the CDF of this distribution. Use your function to plot the distribution where $p=0.70$ where $\theta=0.50$, $\theta=1$, and $\theta=1.50$.**

```{r}
#| label: Alt.CDF
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

paltbinom = function(x, n, p, t){
            i = 0:x
            sum(daltbinom(i, n, p, t)) 
}
#creates CDF as the sum of the PMF

paltbinom(2, 3, 0.7, 1)
paltbinom(2, 3, 0.7, 0.5)
paltbinom(2,3, 0.7, 1.5)

ggdat <- tibble(x = 0:3) |>
rowwise() |> #computes data a row (each x) at a time
mutate(CDF = paltbinom(x,3,0.70,0.5)) #sets theta to 0.5, 
#p to .7, and games to 3

CDF.plot.1 <- ggplot() +
geom_step(data = ggdat, aes(x = x, y = CDF)) +
geom_point(data = ggdat, aes(x = x, y = CDF)) +
geom_hline(yintercept = 0) +
scale_x_continuous("X Where Theta = 0.5", breaks = 0:3) +
ylab("Cumulative Probability") +
theme_bw()

ggdat <- tibble(x = 0:3) |>
  rowwise() |>
  mutate(CDF = paltbinom(x,3,0.70,1))

CDF.plot.2 <- ggplot(ggdat) +
geom_step(data=ggdat, aes(x = x, y = CDF)) +
geom_point(aes(x = x, y = CDF), size = 1) +
geom_hline(yintercept = 0) +
scale_x_continuous("X Where Theta = 1", breaks=0:3) +
ylab("Cumulative Probability") +
theme_bw()

ggdat <- tibble(x = 0:3) |>
  rowwise() |>
  mutate(CDF = paltbinom(x,3,0.70,1.5))

CDF.plot.3 <- ggplot(ggdat) +
geom_step(data=ggdat, aes(x = x, y = CDF)) +
geom_point(aes(x = x, y = CDF), size = 1) +
geom_hline(yintercept = 0) +
scale_x_continuous("X Where Theta = 1.5", breaks=0:3) +
ylab("Cumulative Probability") +
theme_bw()

CDF.plot.1 | CDF.plot.2 | CDF.plot.3
```

## The Winning Probability
**Compute the probability that Player A wins the match using Altham's Multiplicative Binomial Distribution. Compare your result to the Binomial and Negative Binomial results. Compare your results to the Beta-Binomial model, too. Make sure write out your probability statement and work to write it in terms of the distribution function (PMF/CDF) to plot the probability using `ggplot`.**

**Solution**

```{r}
#| label: Winning.Prob
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggdat <- tibble(x = 0:3) |>
 mutate(PMF = daltbinom(x, 3, 0.70, 0.5))

PMF.plot.1 <- ggplot() +
 geom_linerange(data = ggdat,
 aes(x = x, ymin = 0, ymax = PMF)) +
 geom_linerange(data = ggdat |> filter(x >= 2), #highlights where x>= 2
 aes(x = x, ymin = 0, ymax = PMF),
 color = "darkorchid2", linewidth = 2, alpha = 0.50) +
geom_hline(yintercept = 0) +
scale_x_continuous("x", breaks = 0:3) + #breaks for each outcome
ylim(0,1)+
ylab("Probability") +
theme_bw()

ggdat <- tibble(x = 0:3) |>
mutate(PMF = daltbinom(x,3,0.70,1)) # sets theta to 1, 
#all else is same as PMF.plot.1

PMF.plot.2 <- ggplot() +
geom_linerange(data = ggdat,
aes(x = x, ymin = 0, ymax = PMF)) +
geom_linerange(data = ggdat |> filter(x >= 2),
aes(x = x, ymin = 0, ymax = PMF),
color = "darkorchid2", linewidth = 2, alpha = 0.50) +
geom_hline(yintercept = 0) +
scale_x_continuous("x", breaks = 0:3) +
ylim(0,1)+
ylab("Probability") +
theme_bw()

ggdat <- tibble(x = 0:3) |>
mutate(PMF = daltbinom(x,3,0.70,1.5))# sets theta to 1.5, 
#all else is same as PMF.plot.1

PMF.plot.3 <- ggplot() +
geom_linerange(data = ggdat,
aes(x = x, ymin = 0, ymax = PMF)) +
geom_linerange(data = ggdat |> filter(x >= 2),
aes(x = x, ymin = 0, ymax = PMF),
color = "darkorchid2", linewidth = 2, alpha = 0.50) +
geom_hline(yintercept = 0) +
scale_x_continuous("x", breaks = 0:3) +
ylim(0,1)+
ylab("Probability") +
theme_bw()

PMF.plot.1 | PMF.plot.2 | PMF.plot.3
```

$P(X >= 2) = 1 - F(X = 1) = f(X = 2) + f(X = 3)$

```{r}
# theta = 0.5
daltbinom(2, 3, 0.70, 0.5) + daltbinom(3, 3, 0.70, 0.5)
1 - paltbinom(1, 3, 0.70, 0.5)

# theta = 1
daltbinom(2, 3, 0.70, 1) + daltbinom(3, 3, 0.70, 1)
1 - paltbinom(1, 3, 0.70, 1)

# theta = 1.5
daltbinom(2, 3, 0.70, 1.5) + daltbinom(3, 3, 0.70, 1.5)
1 - paltbinom(1, 3, 0.70, 1.5)
```

When theta is 1, the altham and binomial distributions have the same winning probability. When theta is less than one, under the altham distribution there is a greater chance to win all three matches compared to the other distribtuions, while there is a greater chance to only win two matches when theta is greater than one.

# Altham's Multiplicative Beta-Binomial Distribution
You may have wondered -- "Can we do the Beta thing with Altham's version of the Binomial Distribution?" The answer is yes!

## Probability Mass Function
The mass function for this distribution is below. 
$$f_X(x) = \left[\int_{0}^1 dp
\left(\frac{1}{c} {n \choose x} p^x (1-p)^{n-x} \theta^{x(n-x)}\right)\left(\frac{\Gamma(\alpha+\beta)}{\Gamma(\alpha)\Gamma(\beta)} p^{\alpha-1}(1-p)^{\beta-1}\right)\right]I(x \in \{0, 1, ..., n\})$$

**Write a function in `R` called `daltbbinom()` for computing the PMF of this distribution. Use your function to plot the distribution where $\alpha=7$ and $\beta=3$ where $\theta=0.50$, $\theta=1$, and $\theta=1.50$. Comment on the differences in the context of a best two-out-of-three match.**\

**Hint**: You can use the `integrate()` in `R` to compute the integral. Note that this requires the function to be vectorized -- it should be unless you have added any block conditionals. 

**Solution**

```{r}
#| label: Alt.Mult.BBinom
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

daltbbinom = function(x, n, alpha, beta, t){
  integrand = Vectorize(function(p){
  i = 0:n 
  c = sum(choose(n, i)*(p^i)*((1-p)^(n-i)*(t^(i*(n-i)))))

  (1/c)* choose(n, x)*(p^x)*((1-p)^(n-x)*(t^(x*(n-x))))*
  ((gamma(alpha + beta))/(gamma(alpha)*gamma(beta))*(p^(alpha-1))*((1-p)^(beta-1)))})
integrate(integrand, lower = 0, upper = 1)$value 
}
#creates function daltbinom as a function of x, n, alpha, beta, and theta
#This function follows similiarly to the beta binomial function
#However it incorporates alpha and beta, so the gamma function must also be used
# The function is an integral in accordance with the outlined function

ggdat <- tibble(x = 0:3) |>
  rowwise()|> #breaks data up by x value
 mutate(PMF = daltbbinom(x, 3, 7, 3, 0.5))
#x is the independent variable
# 3 games, alpha equals 7, beta equals 3, theta = 0.5

PMF.plot.1 <- ggplot() +
 geom_linerange(data = ggdat,
 aes(x = x, ymin = 0, ymax = PMF)) +
 geom_linerange(data = ggdat |> 
 filter(x >= 2),
 aes(x = x, ymin = 0, ymax = PMF),
 color = "darkorchid2", linewidth = 2, alpha = 0.50) +
  #when x is greater than two, it will be highlighted
geom_hline(yintercept = 0) +
scale_x_continuous("X Where Theta = 0.5", breaks = 0:3) +
ylim(0,1)+
ylab("Probability") +
theme_bw()

ggdat <- tibble(x = 0:3) |>
rowwise()|>
mutate(PMF = daltbbinom(x, 3, 7, 3,1))
#same as PMF.plot.1 but theta = 1

PMF.plot.2 <- ggplot() +
geom_linerange(data = ggdat,
aes(x = x, ymin = 0, ymax = PMF)) +
geom_linerange(data = ggdat |> 
filter(x >= 2),
aes(x = x, ymin = 0, ymax = PMF),
color = "darkorchid2", linewidth = 2, alpha = 0.50) +
geom_hline(yintercept = 0) +
scale_x_continuous("X Where Theta = 1", breaks = 0:3) +
ylim(0,1)+
ylab("Probability") +
theme_bw()

ggdat <- tibble(x = 0:3) |>
rowwise()|>
mutate(PMF = daltbbinom(x,3, 7, 3, 1.5))
#same as PMF.plot.1 but theta = 1.5

PMF.plot.3 <- ggplot() +
geom_linerange(data = ggdat,
aes(x = x, ymin = 0, ymax = PMF)) +
geom_linerange(data = ggdat |> 
filter(x >= 2),
aes(x = x, ymin = 0, ymax = PMF),
color = "darkorchid2", linewidth = 2, alpha = 0.50) +
geom_hline(yintercept = 0) +
scale_x_continuous("X Where Theta = 1.5", breaks = 0:3) +
ylim(0,1)+
ylab("Probability") +
theme_bw()

PMF.plot.1 | PMF.plot.2 | PMF.plot.3
```

Under this distribution, if theta is less than one, then there is a greater liklihood of winning all three matches, if theta is greater than 1 then there is a greater liklihood of winning two matches and losing one. If theta is 1, then there is an equal chance of winning two or three matches.

# A Comparison
Compute the probabilities for all outcomes for Player A.

* Loses 0-2 (LL"L", LL"W")
* Loses 1-2 (WLL, LWL)
* Wins 2-0 (WW"W", WW"L")
* Wins 2-1 (LWW, WLW)

**Note**: You'll need to define these outcomes in terms of the random variable. 

* Binomial:
  * Loses 0-2: $P(X=0)+ \frac{1}{3} P(x=1) = fx(0) + \frac{1}{3} fx(1)$
  * Loses 1-2: $\frac{2}{3} P(X = 1)= \frac{2}{3} fx(1)$
  * Wins 2-0: $P(X=3) + \frac{1}{3} P(X = 2) = fx(3) + \frac{1}{3} fx(2)$
  * Wins 2-1: $\frac{2}{3} P(X = 2) = \frac{2}{3} fx(2)$ 

* Negative Binomial:
  * Wins 2-0: $P(X = 0) = fx(0)$
  * Wins 2-1: $P(X = 1) =fx(1)$
  * Loses: $P(X = 2)=fx(2)$

* Beta-Binomial
  * Loses 0-2: $P(X=0)+ \frac{1}{3} P(x=1) = fx(0) + \frac{1}{3} fx(1)$
  * Loses 1-2: $\frac{2}{3} P(X = 1)= \frac{2}{3} fx(1)$
  * Wins 2-0: $P(X=3) + \frac{1}{3} P(X = 2) = fx(3) + \frac{1}{3} fx(2)$
  * Wins 2-1: $\frac{2}{3} P(X = 2) = \frac{2}{3} fx(2)$ 

* Althman Binomial:
  * Loses 0-2: $P(X=0)+ \frac{1}{3} P(x=1) = fx(0) + \frac{1}{3} fx(1)$
  * Loses 1-2: $\frac{2}{3} P(X = 1)= \frac{2}{3} fx(1)$
  * Wins 2-0: $P(X=3) + \frac{1}{3} P(X = 2) = fx(3) +\frac{1}{3} fx(2)$
  * Wins 2-1: $\frac{2}{3} P(X = 2) = \frac{2}{3} fx(2)$

* Althman Beta-Binomial
  * Loses 0-2: $P(X=0)+ \frac{1}{3} P(x=1) = fx(0) + \frac{1}{3} fx(1)$
  * Loses 1-2: $\frac{2}{3} P(X = 1)= \frac{2}{3} fx(1)$
  * Wins 2-0: $P(X=3) + \frac{1}{3} P(X = 2) = fx(3) +\frac{1}{3} fx(2)$
  * Wins 2-1:  $\frac{2}{3} P(X = 2) = \frac{2}{3} fx(2)$

**Create a bar plot (or column plot) using `ggplot` to show the probability of each outcome under the following conditions:**

* **Binomial**
* **Negative Binomial**\
**Note**: You can compute the lose probabilities the same way as the win probabilities using success probability ($1-p$)
* **Beta-Binomial**
* **Althman Binomial with $\theta=0.50$ and $\theta=1.5$**
* **Althman Beta-Binomial with $\theta=0.50$ and $\theta=1.5$**

```{r}
comparing = tibble(
  outcome = rep(c("Loses 0-2", "Loses 1-2", "Wins 2-0", "Wins 2-1"),times=7),
  #each of these 4 outcomes will be repeated 7 times for each distribution
  distribution = rep(c(
                      "Binomial",
                      "Negative_Binomial",
                      "Beta_Binomial",
                      "Altham(theta = 0.5)",
                      "Altham(theta = 1.5)",
                      "AlthamBeta(theta = 0.5)",
                      "AlthamBeta(theta = 1.5)"), each=4),
 #each distribution is repeated 4 times for each outcome

 #The following is just the probability of each outcome for each distribution
 #provides the data for the graph
  probability= c(
   # Binomial
      dbinom(0,3,0.70)+(1/3)*dbinom(1,3,0.70),
      (2/3)*dbinom(1,3,0.70),
      dbinom(3,3,0.70)+(1/3)*dbinom(2,3,0.70),
      (2/3)*dbinom(2,3,0.70),
    # Negative Binomial
      dnbinom(0,2,0.30), 
      dnbinom(1,2,0.30), 
      dnbinom(0,2,0.70), 
      dnbinom(1,2,0.70),
    # Beta-Binomial
      dbbinom(0,3,7,3)+(1/3)*dbbinom(1,3,7,3),
      (2/3)*dbbinom(1,3,7,3),
      dbbinom(3,3,7,3)+(1/3)*dbbinom(2,3,7,3),
      (2/3)*dbbinom(2,3,7,3),
    # Altham Binomial theta=0.5
      daltbinom(0,3,0.70,0.5)+(1/3)*daltbinom(1,3,0.70,0.5),
      (2/3)*daltbinom(1,3,0.70,0.5),
      daltbinom(3,3,0.70,0.5)+(1/3)*daltbinom(2,3,0.70,0.5),
      (2/3)*daltbinom(2,3,0.70,0.5),
    # Altham Binomial theta=1.5
      daltbinom(0,3,0.70,1.5)+(1/3)*daltbinom(1,3,0.70,1.5),
      (2/3)*daltbinom(1,3,0.70,1.5),
      daltbinom(3,3,0.70,1.5)+(1/3)*daltbinom(2,3,0.70,1.5),
      (2/3)*daltbinom(2,3,0.70,1.5),
    # Altham Beta-Binomial theta=0.5
      daltbbinom(0,3,7,3,0.5)+(1/3)*daltbbinom(1,3,7,3,0.5),
      (2/3)*daltbbinom(1,3,7,3,0.5),
      daltbbinom(3,3,7,3,0.5)+(1/3)*daltbbinom(2,3,7,3,0.5),
      (2/3)*daltbbinom(2,3,7,3,0.5),
    # Altham Beta-Binomial theta=1.5
      daltbbinom(0,3,7,3,1.5)+(1/3)*daltbbinom(1,3,7,3,1.5),
      (2/3)*daltbbinom(1,3,7,3,1.5),
      daltbbinom(3,3,7,3,1.5)+(1/3)*daltbbinom(2,3,7,3,1.5),
      (2/3)*daltbbinom(2,3,7,3,1.5)
  )
)
```
```{r}
#| label: comparing_probabilities
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggplot(comparing) + #comparing ensures its the correct tibble
geom_bar(aes (x = distribution, #distribution on the x-axis
              y = probability), #previously calculated probabilities on y
              stat="identity") +
geom_hline(yintercept = 0)+
theme_bw()+
labs(x = "Distribution Name", 
     y = "Probability") +
ggtitle("Each Outcome's Probability Across Different Distributions") +
facet_wrap(~outcome) + #seperates each bar graph by outcome
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**Comment on the differences in the context of a best two-out-of-three match**

For every probability distribution, there is a greater chance of winning rather than losing. Winning $2-0$ is the greatest liklihood across all probability distributions. For the various Altham Distributions, as theta increases there is a greater chance of losing a match, and therefore as theta approaches 0 there is a greater chance of winning matches consecutively, demonstrating the cluster effect. The Beta Binomial, Negative Biniomial, and Binomial distributions move relatively together, with the negative and binomial distributions being about equal. This makes sense based on the similar math behind both distributions, with the negative binomial distribution counting failures while the binomial distribution counted successes. It makes sense that the beta binomial distribution is similar but not the same as it is still a binomial distribution, but it does not assume a fixed probability of winning a given match. Rather, it highlights the idea of a winning streak (that as you win once you win more).

# Conclusion

I used the tidyverse package [@tidyverse] and the patchwork package [@patchwork] to create graphs, the knitr package [@knitr] to help organize and reference graphs, the extraDistr package [@extraDistr] for the Beta Binomial Distribution, all while using R [@R] as my coding language.